### 2020-05-18 writeup/proof_technical.tex

@@

\subsection{Proof of Theorem ???}

\paragraph{Extension to Many Tasks of the Same Covariates}

In this section we consider the setting with $k$ many that have the same covariates.
Since every task has the same number of data points as well as the same covariance, the only differences between different tasks are their models $\set{\beta_i}_{i=1}^k$.
For this setting, we derive solutions for the multi-task training and the transfer learning setting that match our insights qualitatively from Section \ref{sec_denoise}.

Concretely we will consider the following problem.
\begin{align}
	f(B; W_1, \dots, W_k) = \sum_{i=1}^k \bignorm{X B W_i - Y_i}^2. \label{eq_mtl_same_cov}
\end{align}
By fixing $W_1, W_2, \dots, W_k$, we can derive a closed form solution for $B$ as
\begin{align*}
	\hat{B}(W_1, \dots, W_k) &= (X^{\top}X)^{-1} X^{\top} \bigbrace{\sum_{i=1}^k Y_i W_i^{\top}} (Z Z^{\top})^{-1} \\
	&= \sum_{i=1}^k \bigbrace{\beta_i W_i^{\top}} (ZZ^{\top})^{-1} + (X^{\top}X)^{-1}X^{\top} \bigbrace{\sum_{i=1}^k \varepsilon_i W_i^{\top}} (ZZ^{\top})^{-1}
\end{align*}
where we denote $Z\in\real^{r\times k}$ as the $k$ vectors $W_1, W_2, \dots, W_k$ stacked together.
%Now we switch $\hat{B}$ back into equation \eqref{eq_mtl_same_cov} to
Similar to Section \ref{sec_setup}, we consider minimizing the validation loss over $W_1, W_2, \dots, W_k$ provided with $\hat{B}$.

\paragraph{Jointly minimizing over all tasks.}
Denote by $\varepsilon(W) = \sum_{i=1}^k \varepsilon_i W_i^{\top}$.
We shall decompose the validation loss $\val(\hat{B}; W_1, \dots, W_k)$ into two parts.
The first part is the model shift bias, which is equal to
\begin{align*}
	\sum_{j=1}^k \bigbrace{\bignorm{\Sigma^{1/2}\bigbrace{\sum_{i=1}^k(\beta_i W_i^{\top}) (ZZ^{\top})^{-1} W_j - \beta_j}}^2}
\end{align*}
The second part is the variance, which is equal to
\begin{align*}
	& \sum_{j=1}^k \exarg{\varepsilon_i, \forall i}{\bigbrace{\bigbrace{\sum_{i=1}^k \varepsilon_i W_i^{\top}} (ZZ^{\top})^{-1} W_j}^2} \cdot {\bigtr{\Sigma(X^{\top}X)^{-1}}} \\
	=& \sigma^2 \cdot \bigtr{\Sigma (X^{\top}X)^{-1}}.
\end{align*}
Therefore we shall focus on the minimizer for the model shift bias since the variance part does not depend the weights.
Let us denote $Q\in\real^{k\times k}$ where the $(i,j)$-th entry is equal to $W_i^{\top} (ZZ^{\top})^{-1} W_j$, for any $1\le i, j\le k$.
Let $B^{\star} = [\beta_1, \beta_2, \dots, \beta_k] \in\real^{p \times k}$ denote the true model parameters.
We can now write the validation loss succinctly as follows.
\begin{align*}
	\val(\hat{B}; W_1, \dots, W_k) = \sum_{j=1}^k \bignorm{\Sigma^{1/2} \bigbrace{B^{\star} Q - \beta_j}}^2 + \sigma^2 \cdot \bigtr{\Sigma (X^{\top}X)^{-1}}
\end{align*}
From the above we can solve for $Q$ optimally as \todo{this}.

\paragraph{Minimizing over the target task alone.}
If we only minimize over the the validation loss for the target task, we shall get the following.
\begin{align*}
	\val_j(\hat{B} W_j) = \bignorm{\Sigma^{1/2}\bigbrace{\sum_{i=1}^k W_i^{\top} (ZZ^{\top})^{-1}W_j \beta_i - \beta_j}}^2
	+ \sigma^2 \cdot W_j^{\top} (ZZ^{\top})^{-1} W_j \cdot \bigtr{\Sigma (X^{\top}X)^{-1}}.
\end{align*}

From the above we can obtain three conceptual insights that are consistent with Section \ref{sec_denoise} and \ref{sec_insight}.
\begin{itemize}
	\item The de-noising effect of multi-task learning.
	\item Multi-task training vs single-task training can be either positive or negative.
	\item Transfer learning is better than the other two. And the improvement over multi-task training increases as the model distances become larger.
\end{itemize}

>>>>

%To illustrate the idea, we observe that by using Lemma \ref{lem_minv}, we have that
%\[ \te(\hat{\beta}_t^{\STL}) = \frac{\sigma^2}{n_2 - p}\bigtr{\Sigma_2^{-1}}. \]
%We shall also derive the limit of $\te(\hat{\beta}_t^{\MTL})$.
%\todo{write a brief technical overview}

@@

\subsection{Proof of Theorem \ref{thm_many_tasks}}\label{app_proof_many_tasks}

%In this section we consider the setting with $k$ many that have the same covariates.
%Since every task has the same number of data points as well as the same covariance, the only differences between different tasks are their models $\set{\beta_i}_{i=1}^k$.
%For this setting, we derive solutions for the multi-task training and the transfer learning setting that match our insights qualitatively from Section \ref{sec_denoise}.

For this setting, the problem reduces to the following.
\begin{align}
	f(B; W_1, \dots, W_k) = \sum_{i=1}^k \bignorm{X B W_i - Y_i}^2. \label{eq_mtl_same_cov}
\end{align}
In order to prove Theorem \ref{thm_many_tasks}, we will derive a closed form solution for equation \eqref{eq_mtl_same_cov}.

\begin{proof}[Proof of Theorem \ref{thm_many_tasks}]
	By fixing $W_1, W_2, \dots, W_k$, we can derive a closed form solution for $B$ as
	\begin{align*}
		\hat{B}(W_1, \dots, W_k) &= (X^{\top}X)^{-1} X^{\top} \bigbrace{\sum_{i=1}^k Y_i W_i^{\top}} (Z Z^{\top})^{-1} \\
		&= \sum_{i=1}^k \bigbrace{\beta_i W_i^{\top}} (ZZ^{\top})^{-1} + (X^{\top}X)^{-1}X^{\top} \bigbrace{\sum_{i=1}^k \varepsilon_i W_i^{\top}} (ZZ^{\top})^{-1}
	\end{align*}
	where we denote $Z\in\real^{r\times k}$ as the $k$ vectors $W_1, W_2, \dots, W_k$ stacked together.
	%Now we switch $\hat{B}$ back into equation \eqref{eq_mtl_same_cov} to
	Similar to Section \ref{sec_setup}, we consider minimizing the validation loss over $W_1, W_2, \dots, W_k$ provided with $\hat{B}$.

	Denote by $\varepsilon(W) = \sum_{i=1}^k \varepsilon_i W_i^{\top}$.
	We shall decompose the validation loss $\val(\hat{B}; W_1, \dots, W_k)$ into two parts.
	The first part is the model shift bias, which is equal to
	\begin{align*}
		\sum_{j=1}^k \bigbrace{\bignorm{\Sigma^{1/2}\bigbrace{\sum_{i=1}^k(\beta_i W_i^{\top}) (ZZ^{\top})^{-1} W_j - \beta_j}}^2}
	\end{align*}
	The second part is the variance, which is equal to
	\begin{align*}
		& \sum_{j=1}^k \exarg{\varepsilon_i, \forall i}{\bigbrace{\bigbrace{\sum_{i=1}^k \varepsilon_i W_i^{\top}} (ZZ^{\top})^{-1} W_j}^2} \cdot {\bigtr{\Sigma(X^{\top}X)^{-1}}} \\
		=& \sigma^2 \cdot \bigtr{\Sigma (X^{\top}X)^{-1}}.
	\end{align*}
	Therefore we shall focus on the minimizer for the model shift bias since the variance part does not depend the weights.
	Let us denote $Q = Z^{\top} (ZZ^{\top})^{-1} Z \in\real^{k\times k}$ where the $(i,j)$-th entry is equal to $W_i^{\top} (ZZ^{\top})^{-1} W_j$, for any $1\le i, j\le k$.
	Let $B^{\star} = [\beta_1, \beta_2, \dots, \beta_k] \in\real^{p \times k}$ denote the true model parameters.
	We can now write the validation loss succinctly as follows.
	\begin{align*}
		\val(\hat{B}; W_1, \dots, W_k) = \bignormFro{\Sigma^{1/2} \bigbrace{B^{\star} Q - B^{\star}}}^2 + \sigma^2 \cdot \bigtr{\Sigma (X^{\top}X)^{-1}}
	\end{align*}
	From the above we can solve for $Q$ optimally as $U_{r}U_r^{\top}$.
	Furthermore, we can solve $\hat{\beta}_i^{\MTL}$ as $B^{\star} U_r U_r(i)$.
	Now we get that
	\begin{align*}
		\te(\hat{\beta}_t^{\MTL}) &= \bignorm{\Sigma^{1/2}\bigbrace{\sum_{i=1}^k W_i^{\top} (ZZ^{\top})^{-1}W_j \beta_i - \beta_j}}^2
		+ \sigma^2  W_j^{\top} (ZZ^{\top})^{-1} W_j \cdot \bigtr{\Sigma (X^{\top}X)^{-1}} \\
		&= \bignorm{\Sigma^{1/2} \bigbrace{B^{\star} U_r U_r(i)}}^2 + \sigma^2\norm{U_r(i)}^2 \cdot \bigtr{\Sigma (X^{\top}X)^{-1}}.
	\end{align*}
	By using Lemma \ref{lem_minv}, we conclude the proof.
\end{proof}
%From the above we can obtain three conceptual insights that are consistent with Section \ref{sec_denoise} and \ref{sec_insight}.
%\begin{itemize}
%	\item The de-noising effect of multi-task learning.
%	\item Multi-task training vs single-task training can be either positive or negative.
%	\item Transfer learning is better than the other two. And the improvement over multi-task training increases as the model distances become larger.
%\end{itemize}
